### 2015-05-13 whoop.tex
update in summarisation

Our approach considers, for a given driver, every pair of entry points that can potentially execute concurrently.  For each such pair, we use symbolic verification to check whether it is possible for the pair to race on a shared memory location. We soundly model the effects of additional entry points by treating the driver shared state abstractly: when an entry point reads from the shared state, a nondeterministic value is returned.  Restricting to pairs of entry points, rather than analysing all entry points simultaneously, exploits the fact that data races occur between pairs of threads and allows our analysis to scale to large drivers.  The trade-off is that a quadratic number of entry point pairs must be checked.  In \S\ref{sec:implementation} we discuss optimizations based on device driver domain knowledge to reduce the number of pairs to some extent.

@@

To avoid quantifiers and the associated theorem proving burden, we use instead a \emph{watchdog race-checking instrumentation}, adapted from~\cite{bardsley2014engineering}.  Suppose we are analysing entry points $T$ and $U$, and that after translation into Boogie these entry points share a common memory region, $\mathit{MR}$.  When analyzing $T$ and $U$ for races, we introduce an unconstrained symbolic constant $\mathit{watched}_{\mathit{MR}}$, representing some unspecified index into $\mathit{MR}$; we call this the \emph{watched offset} for $\mathit{MR}$.  We then attempt to prove that it is impossible for $T$ and $U$ to race on $\mathit{MR}$ at index $\mathit{watched}_{\mathit{MR}}$.  If we can succeed in proving this, we know that $T$ and $U$ must be race-free for the \emph{whole} of $\mathit{MR}$, since the watched offset was arbitrary.  This technique of choosing an arbitrary index to analyse for each map manipulated by an entry point pair can be seen as a form of quantifier elimination: rather than asking the underlying theorem prover to reason for all indices of $\mathit{MR}$, in a quantified manner, we eliminate the quantifier in our encoding, and instead ask the theorem prover to reason about a single, but arbitrary, index of $\mathit{MR}$.

@@

\ADComment{We have not given much intuition for what these candidates look like.  Could you give an example candidate, using mathematical notation?  For instance, if entry point $T$ has a procedure $f$ that releases mutex $m$, then when instrumenting $T$ with symbol $i$ (using the notation from \S\ref{sec:symbolicpairwise}) we might guess $m \notin \currentlockset{i}$ as a postcondition for $\instrument{f}{i}$.  It would be ideal if you could be as specific as possible about the candidates we generate.}

The candidate invariants are based on the aforementioned watchdog constants that \whoop instruments for watchdog race-checking. Our candidate generation rules are purely heuristic. The key idea is to automatically generate candidates for each memory region in the granularity of watched accesses. This reduces the false positives from reporting errors if only e.g.\ a field in memory region $\mathit{MR}$ is racing, but the rest of the fields in $\mathit{MR}$ are not racing.

To generate the candidate invariants, \whoop performs an inter-procedural static analysis on the call-graph (in Boogie IVL code) of each entry point, identifying all possible accesses to shared memory locations. To be sound, if \whoop cannot infer information for a specific access, it generates a coarse-grained candidate for the corresponding memory region.

>>>>

Our approach considers, for a given driver, every pair of entry points that can potentially execute concurrently.  For each such pair, we use symbolic verification to check whether it is possible for the pair to race on a shared memory location. We soundly model the effects of additional entry points by treating the driver shared state abstractly: when an entry point reads from the shared state, a nondeterministic value is returned.  Restricting to pairs of entry points, rather than analyzing all entry points simultaneously, exploits the fact that data races occur between pairs of threads and allows our analysis to scale to large drivers.  The trade-off is that a quadratic number of entry point pairs must be checked.  In \S\ref{sec:implementation} we discuss optimizations based on device driver domain knowledge to reduce the number of pairs to some extent.

@@

To avoid quantifiers and the associated theorem proving burden, we use instead a \emph{watchdog race-checking instrumentation}, adapted from~\cite{bardsley2014engineering}.  Suppose we are analyzing entry points $T$ and $U$, and that after translation into Boogie these entry points share a common memory region, $\mathit{MR}$.  When analyzing $T$ and $U$ for races, we introduce an unconstrained symbolic constant $\mathit{watched}_{\mathit{MR}}$, representing some unspecified index into $\mathit{MR}$; we call this the \emph{watched offset} for $\mathit{MR}$.  We then attempt to prove that it is impossible for $T$ and $U$ to race on $\mathit{MR}$ at index $\mathit{watched}_{\mathit{MR}}$.  If we can succeed in proving this, we know that $T$ and $U$ must be race-free for the \emph{whole} of $\mathit{MR}$, since the watched offset was arbitrary.  This technique of choosing an arbitrary index to analyze for each map manipulated by an entry point pair can be seen as a form of quantifier elimination: rather than asking the underlying theorem prover to reason for all indices of $\mathit{MR}$, in a quantified manner, we eliminate the quantifier in our encoding, and instead ask the theorem prover to reason about a single, but arbitrary, index of $\mathit{MR}$.

@@

The candidate invariants are based on the watchdog constants that \whoop instruments for watchdog race-checking. Our candidate generation rules are purely heuristic. The key idea is to automatically generate candidates for each memory region in the granularity of watched accesses. For example, if entry point $T$ has a procedure $f$ that only writes (and not reads) the offset $z$ of memory region $\mathit{MR}$, which is associated with the write set $\hasbeenwritten{\mathit{MR}, i}$ and read set $\hasbeenread{\mathit{MR}, i}$, then when instrumenting $T$ with symbol $i$ (using the notation from \S\ref{sec:symbolicpairwise}), we might guess $\mathit{watched}_{\mathit{MR}} = z \Longrightarrow z \in \hasbeenwritten{\mathit{MR}, i}$ and $\mathit{watched}_{\mathit{MR}} = z \Longrightarrow z \notin \hasbeenread{\mathit{MR}, i}$, as postconditions for $\instrument{f}{i}$. Likewise, if $f$ releases mutex $m$, then we might guess $m \notin \currentlockset{i}$ as a postcondition.

Using watchdog-based candidates reduces the false positives from reporting errors if only e.g.\ a field in memory region $\mathit{MR}$ is racing, but the rest of the fields in $\mathit{MR}$ are not racing. To generate the candidate invariants, \whoop performs an inter-procedural static analysis on the call-graph (in Boogie IVL code) of each entry point, identifying all possible accesses to shared memory locations. To be sound, if \whoop cannot infer information for a specific access, it generates a coarse-grained candidate for the corresponding memory region, e.g.\  $z \in \hasbeenwritten{\mathit{MR}, i}$ and $z \notin \hasbeenread{\mathit{MR}, i}$ for the above example.
